---
title: "Comparing Clustering Methods on 2-Cluster Mixed Data"
output: rmarkdown::github_document
date: '2022-09-08'
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
  echo = TRUE,
  fig.retina = 2,
  message = FALSE,
  warning = FALSE
)

```


The goal of the project was to compare established methods of clustering mixed data (data with both continuous and categorical variables) to `speClusMixed`  our implementation of spectral clustering for mixed data.

The established methods chosen were `KAMILA` and `kprototypes`.

```{r clean-slate, echo=FALSE}

rm(list = ls())
gc()

```


```{r}

# Load KAMILA and kprototypes
library(kamila) # kamila
library(clustMixType) # kproto
library(MixGHD) # ARI

# load functions for mixed data generation

file_to_load <- "/Users/felix.mbuga/Library/Mobile Documents/com~apple~CloudDocs/GitHub/projects/Clustering/functions_gen_2_clusters.R"
source(file_to_load)
ls()

# load function for spectral clustering of mixed data

file_to_load <- "/Users/felix.mbuga/Library/Mobile Documents/com~apple~CloudDocs/GitHub/projects/Clustering/function_speClusMixed.R"
source(file_to_load)
ls()

```

Generate dataset and visualize

```{r}

z <- gen_2_clusters_mixed(seed = 42)
str(z)

pairs(~ ., data = z$data,
      col = z$labels)

```


Cluster dataset with `kamila` and with `kproto` and compare result to `speClusMixed`

```{r}

z_kam <- kamila(Filter(is.numeric, z$data), Filter(is.factor, z$data), numClust = 2, numInit = 100)
print(paste('kamila ARI:', round(ARI(z$labels, z_kam$finalMemb), 3)))

z_kproto <-kproto(z$data, k = 2, nstart = 100, verbose = FALSE)
print(paste('kproto ARI:', round(ARI(z$labels, z_kproto$cluster), 3)))

z_speclus <- speClusMixed(z)
print(paste('speclus ARI:', round(z_speclus$ARI['ARI'], 3)))

```

Another dataset

```{r}

z <- gen_2_clusters_mixed(con_dist = 2, seed = sample(1000, 1))

pairs(~ ., col = z$labels, data = z$data)

z_kam <- kamila(Filter(is.numeric, z$data), Filter(is.factor, z$data), numClust = 2, numInit = 100)
print(paste('kamila ARI:', round(ARI(z$labels, z_kam$finalMemb), 3)))

z_kproto <-kproto(z$data, k = 2, nstart = 100, verbose = FALSE)
print(paste('kproto ARI:', round(ARI(z$labels, z_kproto$cluster), 3)))

z_speclus <- speClusMixed(z)
print(paste('speclus ARI:', round(z_speclus$ARI['ARI'], 3)))

```
Generate multiple (4,800) datasets to compare the three methods on.

Number of continuous-categorical variables 1-3, 2-2, 3-1
Number of levels in categorical variables 3, 5
Degree of overlap in variables lo-lo, lo-hi, hi-lo, hi-hi
Clusters balanced? (points per cluster) 200-200, 320-80

```{r}

##### Step 1: Create the Datasets

## 4800 datasets

# load functions for mixed data generation

file_to_load <- "/Users/felix.mbuga/Library/Mobile Documents/com~apple~CloudDocs/GitHub/projects/Clustering/functions_gen_2_clusters.R"
source(file_to_load)
ls()

# create list with 4 continuous-categorical olap types low-low, low-high, high-low, high-high
data_4800 <- vector(mode = 'list', length = 4) # 4 overlap profiles, 20 datasets each, 80 datasets total
names(data_4800) <- c('low_low', 'low_high', 'high_low', 'high_high')
#str(data_4800)

# for each olap type, number of continuous-categorical variables: 3-1, 2-2, 1-3
for (i in 1:length(data_4800)) {
	#print(i)
	#str(i)
	data_4800[[i]] <- vector(mode = 'list', length = 3)
	names(data_4800[[i]]) <- c('con_3_cat_1', 'con_2_cat_2', 'con_1_cat_3')
}
#str(data_4800)

# for each con-cat ratio, 2 categorical level types: 3 & 5
for (i in 1:4) { # 4 olap types
	for (j in 1:3) { # 3 con-cat ratios
		data_4800[[i]][[j]] <- vector(mode = 'list', length = 2)
		names(data_4800[[i]][[j]]) <- c('3_cat_levs', '5_cat_levs')
	}
}
#str(data_4800)

# for each cat level type, 2 balance levels: 200/200 & 320/80
for (i in 1:4) { # 4 olap types
	for (j in 1:3) { # 3 con-cat ratios
		for (k in 1:2) { # 2 cat-lev types
			data_4800[[i]][[j]][[k]] <- vector(mode = 'list', length = 2)
			names(data_4800[[i]][[j]][[k]]) <- c('200_200', '320_80')
		}
	}
}
#str(data_4800)

# under each dataset type, have 100 lists, 1 for each of 100 datasets
for (i in 1:4) { # 4 olap types
	for (j in 1:3) { # 3 con-cat ratios
		for (k in 1:2) { # 2 cat-lev types
			for (l in 1:2) { # 2 balance levels
				data_4800[[i]][[j]][[k]][[l]] <- vector(mode = 'list', length = 100)
				names(data_4800[[i]][[j]][[k]][[l]]) <- as.character(1:100)
			}		
		}
	}
}
#str(data_4800)


# under each dataset, store 4 items
	# - the dataset & cluster labels
	# - the kamila output
	# - the k-prototypes output
	# - the speclus output
for (i in 1:4) { # 4 olap types
	for (j in 1:3) { # 3 con-cat ratios
		for (k in 1:2) { # 2 cat-lev types
			for (l in 1:2) { # 2 balance levels
				for (m in 1:100) { # 100 datasets
					data_4800[[i]][[j]][[k]][[l]][[m]] <- vector(mode = 'list', length = 4)
					names(data_4800[[i]][[j]][[k]][[l]][[m]]) <- c('dataset', 'kamila', 'kproto', 'speclus')
				}			
			}		
		}
	}
}
#str(data_4800)

data_4800[[i]][[j]][[k]][[l]][[m]]	# list of 4 items: c('dataset', 'kamila', 'kproto', 'speclus')

##### %%% next step: make the 4800 datasets to populate data_4800

# make the 4800 datasets; calculate kamila & kproto ARI
set.seed(4800)

con_dists <- c(2, 2, 1, 1)
cat_overlaps <- c("low", "high", "low", "high")
(olaps <- data.frame(con_dists, cat_overlaps))

(con_cat_ratio <- matrix(c(3, 1, 2, 2, 1, 3), nrow = 3, byrow = TRUE))
(cat_lev_types <- c(3, 5))
(bal_levs <- matrix(c(200, 200, 320, 80), nrow = 2, byrow=TRUE))


count <- 0
for (i in 1:4) { # 4 olap types
	for (j in 1:3) { # 3 con-cat ratios
		for (k in 1:2) { # 2 cat-lev types
			for (l in 1:2) { # 2 balance levels
				for (m in 1:100) { # 100 datasets
					z <- data_4800[[i]][[j]][[k]][[l]][[m]]$dataset <- 
					  gen_2_clusters_mixed(cluster_points = c(bal_levs[l, 1], bal_levs[l, 2]), 
					                       con_vars =  con_cat_ratio[j, 1],
					                       con_dist =  olaps[i, 1], 
					                       seed = count,
					                       cat_vars = con_cat_ratio[j, 2], 
					                       cat_overlap = olaps[i, 2])
					count = count + 1
					#print(count)
				}			
			}		
		}
	}
}


```
 
```{r}

# str(data_4800)

```

```{r}

ptm <- proc.time()

for (i in 1:4) { # 4 olap types
	for (j in 1:3) { # 3 con-cat ratios
		for (k in 1:2) { # 2 cat-lev types
			for (l in 1:2) { # 2 balance levels
				for (m in 1:100) { # 100 datasets
					print(c(i, j, k, l, m))
					cat('\n')
					z <- data_4800[[i]][[j]][[k]][[l]][[m]]$dataset
					
					z_kam <- data_4800[[i]][[j]][[k]][[l]][[m]]$kamila$z_kam <- kamila(Filter(is.numeric, z$data), Filter(is.factor, z$data), 
																			numClust = 2, numInit = 100)
					#print(paste('kamila ARI:', round(ARI(z$lbl, z_kam$finalMemb), 3)))
					data_4800[[i]][[j]][[k]][[l]][[m]]$kamila$ARI <- round(ARI(z$labels, z_kam$finalMemb), 3)
					
					z_kproto <- data_4800[[i]][[j]][[k]][[l]][[m]]$kproto$z_kproto <- kproto(z$data, k = 2, nstart = 100, verbose = FALSE)
					#print(paste('kproto ARI:', round(ARI(z$lbl, z_kproto$cluster), 3)))
					data_4800[[i]][[j]][[k]][[l]][[m]]$kproto$ARI <- round(ARI(z$labels, z_kproto$cluster), 3)
					
					try({
					  
					  z_speclus <- data_4800[[i]][[j]][[k]][[l]][[m]]$speclus$z_speclus <- speClusMixed(z)
					
					  }) # end try
					
					#count <- count + 1
					#print(count)
					#cat('\n')
				}			
			}		
		}
	}
}

proc.time() - ptm

```

Dataset 1-1-1-1-1 crashes SpeClusMixed with error message:

```

Error in kmeans(yi, centers, iterations, nstart = starts) : 
  more cluster centers than distinct data points.
In addition: Warning message:
In do.call(.Call, args = dot_call_args) :
  only 0 eigenvalue(s) converged, less than k = 2

```

All other datasets run fine.

---

Adding `try` allowed the above to run on the HPC. Took a little over 22 hours to run to completion.

```{r}

# clean slate

rm(list = ls())
gc()

# load RData with result

file_name <- '/Users/felix.mbuga/Library/Mobile Documents/com~apple~CloudDocs/Random_R/data_4800_2022_10_22_09_14_56.RData'
ne <- new.env()
load(file = file_name, env=ne)
ls(env=ne)

```

```{r}

data_4800 <- ne$data_4800
rm(ne, file_name)
ls()

```

Make dataframe with all ARIs for boxplots

```{r Make dataframe with all ARIs for boxplots}

# see 200401 scratch.R and 2103 scratch.R

# make dataframe with all ARIs for boxplots
overlap <- character(14400)
overlap_ <- c('lo-lo', 'lo-hi', 'hi-lo', 'hi-hi')
#overlap <- rep(overlap_, each = 1200)
# overlap <- rep(overlap_, each = 12)

con_cat_vars <- character(14400)
con_cat_vars_ <- c('3-num-1-cat', '2-num-2-cat', '1-num-3-cat')
# con_cat_vars <- rep(con_cat_vars_, each = 4, times = 4)
# data.frame(overlap, con_cat_vars)

cat_levs <- integer(14400)
cat_levs_ <- c(3, 5)
# cat_levs <- rep(cat_levs_, each = 2, times = 12)
# data.frame(overlap, con_cat_vars, cat_levs)

cluster_sizes <- character(14400)
cluster_sizes_ <- c('200-200', '320-80')
# cluster_sizes <- rep(cluster_sizes_, each = 1, times = 24)
# data.frame(overlap, con_cat_vars, cat_levs, cluster_sizes)


dataset_index <- integer(14400)
dataset_index_ <- 1:100
# dataset_index <- rep(dataset_index_, each = 1, times = 48)
# data.frame(overlap, con_cat_vars, cat_levs, cluster_sizes, dataset_index)

method <- character(14400)
method_ <- c('kpr', 'kam', 'spc')
# method <- rep(method_, each = 1, times = 48)
# data.frame(overlap, con_cat_vars, cat_levs, cluster_sizes, dataset_index, method)

ARI_x <- numeric(14400)

n <- 1
for (i in 1:4) { # 4 olap types
	for (j in 1:3) { # 3 con-cat ratios
		for (k in 1:2) { # 2 cat-lev types
			for (l in 1:2) { # 2 balance levels
				for (m in 1:100) { # 100 datasets
					#print(c(i, j, k, l, m))
					overlap[(3*n-2):(3*n)] <- overlap_[i]
					con_cat_vars[(3*n-2):(3*n)] <- con_cat_vars_[j]
					cat_levs[(3*n-2):(3*n)] <- cat_levs_[k]
					cluster_sizes[(3*n-2):(3*n)] <- cluster_sizes_[l]
					dataset_index[(3*n-2):(3*n)] <- dataset_index_[m]
					method[(3*n-2):(3*n)] <- c('kpr', 'kam', 'spc')
					
					kpr_ARI <- data_4800[[i]][[j]][[k]][[l]][[m]]$kproto$ARI
					kam_ARI <- data_4800[[i]][[j]][[k]][[l]][[m]]$kamila$ARI
					spc_ARI <- data_4800[[i]][[j]][[k]][[l]][[m]]$speclus$z_speclus$ARI['ARI']
					
					ARI_x[(3*n-2):(3*n)] <- c(kpr_ARI, kam_ARI, spc_ARI)
					
					#cat('\n')
					
					#z_speclus <- data_4800[[i]][[j]][[k]][[l]][[m]]$speclus$z_speclus <- speclus_v(z)
					
					#count <- count + 1
					#print(n)
					#cat('\n')
					n <- n + 1
				}			
			}		
		}
	}
}

data_4800_ARI <- data.frame(overlap, con_cat_vars, cat_levs, 
                            cluster_sizes, dataset_index, method, ARI_x)

head(data_4800_ARI)

```

Convert relevant variables to factor

```{r}

str(data_4800_ARI)

data_4800_ARI$cat_levs <- as.factor(data_4800_ARI$cat_levs)
data_4800_ARI$dataset_index <- as.factor(data_4800_ARI$dataset_index)

head(data_4800_ARI)


```

Boxplots!

```{r}

boxplot(ARI_x ~ ., data = data_4800_ARI[-5]) # too much info on 1 plot! split up into lo-lo, lo-hi, hi-lo, hi-hi


```

Too much info on 1 plot! split up into lo-lo, lo-hi, hi-lo, hi-hi

```{r lo-lo boxplot}

data_4800_ARI_lo_lo <- data_4800_ARI[data_4800_ARI$overlap == 'lo-lo', -5]
data_4800_ARI_lo_lo$overlap <- factor(data_4800_ARI_lo_lo$overlap)
x <- boxplot(ARI_x ~ ., data = data_4800_ARI_lo_lo, col = 1:12)
x$names
legend('bottomleft', data_4800_ARI_lo_lo$method, fill = c(1, 2, 3))


```

Unstack

```{r lo-lo unstack}


aggregate(ARI_x ~ ., data = data_4800_ARI_lo_lo, median)
unstack(aggregate(ARI_x ~ ., data = data_4800_ARI_lo_lo, median), ARI_x ~ method)

x <- aggregate(ARI_x ~ ., data = data_4800_ARI[data_4800_ARI$overlap == 'lo-lo', -5], function(x) round(median(x), 3))
y <- unstack(aggregate(ARI_x ~ ., data = data_4800_ARI[data_4800_ARI$overlap == 'lo-lo', -5], function(x) round(median(x), 3)), ARI_x ~ method)
(data_4800_ARI_lo_lo_median <- cbind(subset(x[1:12, ], select = -c(ARI_x, method)), y))

```

```{r lo-hi}

x <- aggregate(ARI_x ~ ., data = data_4800_ARI[data_4800_ARI$overlap == 'lo-hi', -5], function(x) round(median(x), 3))
y <- unstack(aggregate(ARI_x ~ ., data = data_4800_ARI[data_4800_ARI$overlap == 'lo-hi', -5], function(x) round(median(x), 3)), ARI_x ~ method)
(data_4800_ARI_lo_hi_median <- cbind(subset(x[1:12, ], select = -c(ARI_x, method)), y))

```

```{r hi-lo}

x <- aggregate(ARI_x ~ ., data = data_4800_ARI[data_4800_ARI$overlap == 'hi-lo', -5], function(x) round(median(x), 3))
y <- unstack(aggregate(ARI_x ~ ., data = data_4800_ARI[data_4800_ARI$overlap == 'hi-lo', -5], function(x) round(median(x), 3)), ARI_x ~ method)
(data_4800_ARI_hi_lo_median <- cbind(subset(x[1:12, ], select = -c(ARI_x, method)), y))

```

```{r hi-hi}

x <- aggregate(ARI_x ~ ., data = data_4800_ARI[data_4800_ARI$overlap == 'hi-hi', -5], function(x) round(median(x), 3))
y <- unstack(aggregate(ARI_x ~ ., data = data_4800_ARI[data_4800_ARI$overlap == 'hi-hi', -5], function(x) round(median(x), 3)), ARI_x ~ method)
(data_4800_ARI_hi_hi_median <- cbind(subset(x[1:12, ], select = -c(ARI_x, method)), y))

```

## spc_ARI vs spc_ARI_best_possible


```{r}

# load data_4800 RData

file_name <- '/Users/felix.mbuga/Library/Mobile Documents/com~apple~CloudDocs/Random_R/data_4800_2022_10_22_09_14_56.RData'
ne <- new.env()
load(file = file_name, env=ne)
data_4800 <- ne$data_4800
rm(ne, file_name)


spc_ARI <- spc_ARI_best_possible <- cat_wgt <- cat_wgt_best_pos <- discrepancy <- kam_less_spc <- numeric(4800)

n <- 1
for (i in 1:4) { # 4 olap types
	for (j in 1:3) { # 3 con-cat ratios
		for (k in 1:2) { # 2 cat-lev types
			for (l in 1:2) { # 2 balance levels
				for (m in 1:100) { # 100 datasets

					spc_ARI[n] <- data_4800[[i]][[j]][[k]][[l]][[m]]$speclus$z_speclus$ARI['ARI']
					cat_wgt[n] <- data_4800[[i]][[j]][[k]][[l]][[m]]$speclus$z_speclus$ARI['cat_wgt']
					spc_ARI_best_possible[n] <- data_4800[[i]][[j]][[k]][[l]][[m]]$speclus$z_speclus$ARI_best_possible['ARI']
					cat_wgt_best_pos[n] <- data_4800[[i]][[j]][[k]][[l]][[m]]$speclus$z_speclus$ARI_best_possible['cat_wgt']
					discrepancy[n] <- spc_ARI_best_possible[n] - spc_ARI[n]
					kam_less_spc[n] <- data_4800[[i]][[j]][[k]][[l]][[m]]$kamila$ARI - spc_ARI[n]

					
					n <- n + 1
				}			
			}		
		}
	}
}

x <- cbind(spc_ARI, spc_ARI_best_possible, cat_wgt, cat_wgt_best_pos, discrepancy, kam_less_spc)
x <- round(x, 4)

head(x)

```

how different are spc_ARI vs spc_ARI_best_possible?

```{r}

summary(x)
cat('\n')
sum(x[, 'discrepancy'] >= 0.1)
cat('\n')
summary(x[x[, 'kam_less_spc'] > 0, 'kam_less_spc'])

```

sort x by largest discrepancy

```{r}

(x_ordered <- x[order(x[, 'discrepancy'], decreasing = TRUE), ])

```

```{r}

summary(lm(kam_less_spc ~ discrepancy, 
           data = data.frame(x[x[, 'kam_less_spc'] > 0, ])))

```

```{r}

(x_ordered_kam_wins <- x_ordered[x_ordered[, 'kam_less_spc'] > 0, ])

```

```{r}

summary(x_ordered_kam_wins)
cat('\n')
quantile(x_ordered_kam_wins[, 'cat_wgt'], probs = seq(0, 1, 0.1))

```

```{r}

(x_ordered_kam_wins_discr <- x_ordered_kam_wins[x_ordered_kam_wins[, 'discrepancy'] > 0, ])

```

```{r}

(x_ordered_kam_wins_discr_0p3 <- x_ordered_kam_wins[x_ordered_kam_wins[, 'discrepancy'] > 0.3, ])
tail(x_ordered_kam_wins_discr_0p3)

```

Algo unstable at
- very low cat wgt
- very high cat wgt
- some intermediate values e.g. see tail
  - maybe before accepting a value of cat_wgt, check if result(??) changes much +/- 0.01 cat_wgt?
- make dataset harder for kam?

```{r}

sum(x_ordered_kam_wins_discr_0p3[, 'cat_wgt'] > 0.9)

```

```{r}

# clean slate

rm(list = ls())
gc()

# load RData with result

file_name <- '/Users/felix.mbuga/Library/Mobile Documents/com~apple~CloudDocs/Random_R/data_4800_2022_10_22_09_14_56.RData'
ne <- new.env()
load(file = file_name, env=ne)

data_4800 <- ne$data_4800
rm(ne, file_name)
ls()

```


Make dataframe with all ARIs for boxplots

```{r Make dataframe with all ARIs for boxplots}

# see 200401 scratch.R and 2103 scratch.R

# make dataframe with all ARIs for boxplots
overlap <- character(14400)
overlap_ <- c('lo-lo', 'lo-hi', 'hi-lo', 'hi-hi')
#overlap <- rep(overlap_, each = 1200)
# overlap <- rep(overlap_, each = 12)

con_cat_vars <- character(14400)
con_cat_vars_ <- c('3-num-1-cat', '2-num-2-cat', '1-num-3-cat')
# con_cat_vars <- rep(con_cat_vars_, each = 4, times = 4)
# data.frame(overlap, con_cat_vars)

cat_levs <- integer(14400)
cat_levs_ <- c(3, 5)
# cat_levs <- rep(cat_levs_, each = 2, times = 12)
# data.frame(overlap, con_cat_vars, cat_levs)

cluster_sizes <- character(14400)
cluster_sizes_ <- c('200-200', '320-80')
# cluster_sizes <- rep(cluster_sizes_, each = 1, times = 24)
# data.frame(overlap, con_cat_vars, cat_levs, cluster_sizes)


dataset_index <- integer(14400)
dataset_index_ <- 1:100
# dataset_index <- rep(dataset_index_, each = 1, times = 48)
# data.frame(overlap, con_cat_vars, cat_levs, cluster_sizes, dataset_index)

method <- character(14400)
method_ <- c('kpr', 'kam', 'spc')
# method <- rep(method_, each = 1, times = 48)
# data.frame(overlap, con_cat_vars, cat_levs, cluster_sizes, dataset_index, method)

ARI_x <- numeric(14400)

n <- 1
for (i in 1:4) { # 4 olap types
	for (j in 1:3) { # 3 con-cat ratios
		for (k in 1:2) { # 2 cat-lev types
			for (l in 1:2) { # 2 balance levels
				for (m in 1:100) { # 100 datasets
					#print(c(i, j, k, l, m))
					overlap[(3*n-2):(3*n)] <- overlap_[i]
					con_cat_vars[(3*n-2):(3*n)] <- con_cat_vars_[j]
					cat_levs[(3*n-2):(3*n)] <- cat_levs_[k]
					cluster_sizes[(3*n-2):(3*n)] <- cluster_sizes_[l]
					dataset_index[(3*n-2):(3*n)] <- dataset_index_[m]
					method[(3*n-2):(3*n)] <- c('kpr', 'kam', 'spc')
					
					kpr_ARI <- data_4800[[i]][[j]][[k]][[l]][[m]]$kproto$ARI
					kam_ARI <- data_4800[[i]][[j]][[k]][[l]][[m]]$kamila$ARI
					spc_ARI <- data_4800[[i]][[j]][[k]][[l]][[m]]$speclus$z_speclus$ARI['ARI']
					
					ARI_x[(3*n-2):(3*n)] <- c(kpr_ARI, kam_ARI, spc_ARI)
					
					n <- n + 1
				}			
			}		
		}
	}
}

data_4800_ARI <- data.frame(overlap, con_cat_vars, cat_levs, 
                            cluster_sizes, dataset_index, method, ARI_x)

# Convert relevant variables to factor

data_4800_ARI$cat_levs <- as.factor(data_4800_ARI$cat_levs)
data_4800_ARI$dataset_index <- as.factor(data_4800_ARI$dataset_index)

str(data_4800_ARI)

head(data_4800_ARI)

```

unstack ARI_x by method

```{r}

head(unstack(data_4800_ARI, ARI_x ~ method))

```



unstack ARI_x by method and keep rest of columns

```{r}

data_4800_ARI_unstack <- (data.frame(
  unique(data_4800_ARI[1:5]), 
  round(unstack(data_4800_ARI, ARI_x ~ method), 3)
))

head(data_4800_ARI_unstack)

```

in which of the 4 overlap types does spc perform worst compared to kam?

```{r}

# kam - spc
data_4800_ARI_unstack['kam_less_spc'] <- data_4800_ARI_unstack['kam'] - data_4800_ARI_unstack['spc']
head(data_4800_ARI_unstack)

# median kam_less_spc by overlap

head(aggregate(kam_less_spc ~ ., data = data_4800_ARI_unstack[c(1:4, 9)], 
               function(x) round(median(x), 3)))

```

```{r}

x <- aggregate(kam_less_spc ~ ., data = data_4800_ARI_unstack[c(1:4, 9)], 
               function(x) round(median(x), 3))

(x_ordered <- x[order(x[, 'kam_less_spc'], decreasing = TRUE), ])

```

* spc performs worst (compared to kam) in lo-hi and hi-hi
* this is not what was seen in the paper
* investigate first-row datasets

```{r}

head(data_4800_ARI_unstack[(data_4800_ARI_unstack$overlap == 'lo-hi') & 
                             (data_4800_ARI_unstack$con_cat_vars == '1-num-3-cat') & 
                             (data_4800_ARI_unstack$cat_levs == 5) & 
                             (data_4800_ARI_unstack$cluster_sizes == '200-200') & 
                             (data_4800_ARI_unstack$kam_less_spc > 0), -c(1:4)], 
      n = 100)

```

also compare spc_ARI_best_possible

```{r}

spc_ARI <- spc_ARI_best_possible <- cat_wgt <- cat_wgt_best_pos <- discrepancy <- kam_less_spc <- numeric(4800)

n <- 1
for (i in 1:4) { # 4 olap types
	for (j in 1:3) { # 3 con-cat ratios
		for (k in 1:2) { # 2 cat-lev types
			for (l in 1:2) { # 2 balance levels
				for (m in 1:100) { # 100 datasets

					spc_ARI[n] <- data_4800[[i]][[j]][[k]][[l]][[m]]$speclus$z_speclus$ARI['ARI']
					cat_wgt[n] <- data_4800[[i]][[j]][[k]][[l]][[m]]$speclus$z_speclus$ARI['cat_wgt']
					spc_ARI_best_possible[n] <- data_4800[[i]][[j]][[k]][[l]][[m]]$speclus$z_speclus$ARI_best_possible['ARI']
					cat_wgt_best_pos[n] <- data_4800[[i]][[j]][[k]][[l]][[m]]$speclus$z_speclus$ARI_best_possible['cat_wgt']
					discrepancy[n] <- spc_ARI_best_possible[n] - spc_ARI[n]
					kam_less_spc[n] <- data_4800[[i]][[j]][[k]][[l]][[m]]$kamila$ARI - spc_ARI[n]

					
					n <- n + 1
				}			
			}		
		}
	}
}

x <- cbind(spc_ARI, spc_ARI_best_possible, cat_wgt, cat_wgt_best_pos, discrepancy, kam_less_spc)
x <- round(x, 4)

head(x)

```


```{r}

head(cbind(data_4800_ARI_unstack, round(spc_ARI_best_possible, 3)), n = 10)

```

```{r}

x <- cbind(data_4800_ARI_unstack, round(spc_ARI_best_possible, 3))
  
str(x)  

head(x)
```

Inspect `lo-hi	1-num-3-cat	5	200-200` datasets based on ~4 cells above

```{r}

x <- x[order(x[, 'kam_less_spc'], decreasing = TRUE), ]

head(x[(x$overlap == 'lo-hi') & 
                             (x$con_cat_vars == '1-num-3-cat') & 
                             (x$cat_levs == 5) & 
                             (x$cluster_sizes == '200-200') & 
                             (x$kam_less_spc > 0), -c(1:4)], 
      n = 100)

```

start with datasets where spc_ARI_best_possible > kam??

```{r}

y <- x[(x$overlap == 'lo-hi') & 
                             (x$con_cat_vars == '1-num-3-cat') & 
                             (x$cat_levs == 5) & 
                             (x$cluster_sizes == '200-200') & 
                             (x$kam_less_spc > 0), -c(1:4)]
head(y)

```

```{r}

y[y[, 6] > y[, 2], ]

```

- start with `lo-hi	1-num-3-cat	5	200-200` dataset 13

RUN ALL BELOW LINE 579


```{r, eval=FALSE}

x <- aggregate(ARI_x ~ ., data = data_4800_ARI[data_4800_ARI$overlap == 'hi-hi', -5], function(x) round(median(x), 3))
y <- unstack(aggregate(ARI_x ~ ., data = data_4800_ARI[data_4800_ARI$overlap == 'hi-hi', -5], function(x) round(median(x), 3)), ARI_x ~ method)
(data_4800_ARI_hi_hi_median <- cbind(subset(x[1:12, ], select = -c(ARI_x, method)), y))

```




---

Session info

```{r}

R.Version()
R.version
version
sessionInfo()

```
